Long-form recording of infant body position in the home using wearable inertial sensors

### 2023-03-27 manuscript.Rmd
revised whole paper

Three online repositories contain openly shared data, materials, and analysis code. A Databrary repository (https://nyu.databrary.org/volume/1580) includes an exemplar participant's recording session, with the raw video data files, the Datavyu annotations of those video files, a log file with machine-readable synchronization points and nap/diaper change times, and accelerometer and gyroscope data for each of the 4 sensors. A CodeOcean capsule (LINK) hosts an R environment with the exemplar participant's data and source code to: 1) synchronize IMU and video annotations, 2) calculate windowed motion features for their data, and 3) train and test the body position classifier using an "individual model". The chosen participant's classification accuracy was at the median level. Because of the overall size of the full data set and computational power/time required to synchronize and create windowed datasets, it would not be feasible to create a reproducible environment to process the entire dataset. However, in a second CodeOcean capsule (LINK) we share the entire dataset of windowed motion features with corresponding body position codes used to validate the method (described in the results below), along with the processed datafiles used for each set of analyses. This reproducible manuscript created in *papaja* [@papaja] can be executed in the CodeOcean environment for full computational transparency and can be downloaded from Github (LINK).
<!-- FINAL LINKS -->

@@

We report three sets of results based on `r length(unique(ppts$id_uni))` full-day testing sessions resulting in a total of `r round(sum(duration$duration_hours))` hours of movement recording. First, we focus on the accuracy of group and individual models trained and validated from data during the proximal comparison period. This period is unique in having the highest degree of temporal synchrony between video and motion data, allowing us to assess the accuracy of individual body position events. Second, we applied models trained from data during the proximal comparison to predict infant behavior during the distal comparison, when coarser synchrony between video and motion data was available. This distal comparison is critical because it provides the first-ever test of how well infant body classification models predict behavior beyond the initial, supervised period. Data from the first two sets of analyses confirm that body position classification models are accurate across a long period of time, suggesting that data from the entire day will be valid. In the third set of results we examine the data quality of full-day recordings and test whether full-day recordings capture well-established age differences in body position experiences.

@@

The first set of analyses use data from the proximal comparison to determine the "best case" accuracy of the models, training and testing on similar types of data. The high degree of temporal synchronization between video and motion data during this period makes it possible to link human-coded body position annotations to each 4-s window of motion data, providing ground truth data for model training and testing. As in past work [@FranchakScott2021], we compared two types of models: *group models* and *individual models*. To assess the accuracy of body position classification for each recording session, we reserved the last 25% of a session's proximal comparison data as the testing set. The testing set was never used as training data. The two types of models differed based on what data were used during training, but the same random forest modeling procedure was used in both cases. A group model was trained for each session using an aggregated dataset containing the first 75% of all **other** sessions' proximal comparison data. This leave-one-out cross-validation tested the generalization of the model to a recording session that was not used at all in the training set. In contrast, individual models were trained using the first 75% of a recording session, then tested against the (temporally) last 25% of the same session. The individual model tests whether earlier training data generalize to later testing data within an individual participant's recording. In both cases, the testing set was the same, allowing for direct comparison between the two modeling approaches. Figure \@ref(fig:metrics), Table \@ref(tab:metrics-table), and Table \@ref(tab:metrics-by-class) summarize the performance of group and individual models using standard metrics for classification.

@@

Overall accuracy (Figure \@ref(fig:metrics)A) represents the proportion of 4-s windows in the testing set in which the model prediction matched the human annotation of body position. Overall accuracy for group models  (*M* = `r metrics_lookup["Overall Accuracy", "mean_Group"]`) was slightly lower than accuracy for individual models (*M* = `r metrics_lookup["Overall Accuracy", "mean_Individual"]`). Although overall accuracy from our semi-supervised, in-home data collection did not match the near-perfect accuracy  (.95-.98) found in prior in-lab studies [@FranchakScott2021; @AiraksinenRasanen2020], both models approached the level of agreement found between two human coders (*M* = .906). Most likely, lower accuracy in the current study results from the more variable and complex behavior observed in an unsupervised setting rather than from a difference in the quality of the classification model. Visual inspection of Figure \@ref(fig:metrics)A shows that accuracy values were heavily skewed, with most approaching perfect accuracy but a few participants with very poor accuracy. Looking at the median performance suggests that the difference between models was not as wide for the typical participant (group median accuracy = `r metrics_lookup["Overall Accuracy", "median_Group"]`; individual median accuracy = `r metrics_lookup["Overall Accuracy", "median_Individual"]`). Indeed, it is notable that the worst-case accuracy from group models (minimum = `r metrics_lookup["Overall Accuracy", "min_Group"]`) was considerably lower than in individual models (minimum = `r metrics_lookup["Overall Accuracy", "min_Individual"]`). Possibly, individual models accounted for idiosyncrasies in behavior or inconsistencies in sensor placement for those sessions that group models could not generalize to.

### Cohen's Kappa 
Strong overall accuracy can be misleading when class prevalence is unbalanced. For example, most infants spent more time sitting than prone, so good performance for sitting classification could mask poorer performance in prone classification. Accordingly, we report Cohen's kappa, a commonly-used metric that penalizes missing rare events (Figure \@ref(fig:metrics)B), and provide classification metrics for each individual body position (Table \@ref(tab:metrics-by-class)) to account for imbalance in body position rates within and between individuals. Similar to overall accuracy, kappa values were strong for both model types with group kappas (*M* = `r metrics_lookup["Kappa", "mean_Group"]`) somewhat worse compared with individual kappas (*M* = `r metrics_lookup["Kappa", "mean_Individual"]`). Guidelines for interpreting kappa statistics [@LandisKoch1977] consider 0.81–1.00 "Almost Perfect,” 0.61–0.80 “Substantial,” 0.41–0.60 “Moderate,” 0.21– 0.40 “Fair,” and 0–0.20 “Slight to Poor”, indicating that agreement for most group and individual model predictions fell in the Substantial to Almost Perfect range.

@@

Figure \@ref(fig:part2overall) and Table \@ref(tab:pt2overalltable) summarize the overall agreement during the distal comparison period. For each session, we calculated the actual prevalence of each body position as a percentage of the time that the infant was on video and awake using human annotated body position (x-axis on Figure \@ref(fig:part2overall)). Predicted prevalence was calculated the same way for group and individual model predictions, omitting the off-camera and nap periods to make a direct comparison. Overall---across participants and across body position classes---agreement was strong: The correlation between group model predictions and human-coded prevalence was *r* = `r group_agree_cor`, and the correlation between individual model predictions and human-coded prevalence was *r* = `r split_agree_cor`. As in the proximal comparison, agreement varied somewhat between body positions; in particular, agreement for held was poor. Unlike in the proximal period, some body positions were better predicted by group models and others by individual models.

Visual inspection of Figure \@ref(fig:part2overall) indicated two extreme outliers, which we marked by a gray square and a gray diamond. We closely investigated each of the two outliers to understand why their agreement was below the standard of the other sessions. The "gray square" outlier had significant confusion between sitting and supine classification. Reviewing the video indicated that this participant spent a long period of time in a seating device that was reclined almost exactly at 45 degrees, making it difficult to determine if the infant was sitting or supine. The infant also spent a long time in the mother's arms in an ambiguous supine/sitting position that was hard to classify. This participant's proximal accuracy was also poor because similar ambiguities appeared during the initial recording. In contrast, the "gray diamond outlier" had strong proximal accuracy, with confusion only arising in the distal period between upright and held categories. Reviewing the video showed that all disagreements occurred when the infant was in a baby walker; human coders scored this as "upright" but the models predicted it as "held". Most likely, the infant's movements in the baby walker were more similar to how a baby moved while carried, and unlike how most infants moved while walking upright.

@@

Collecting real-time data that can be synchronized to other sources is a key goal in our development of long-form body position classification. Although overall aggregate agreement in the distal comparison was strong, it is important to show that similarly strong agreement is found within a shorter timescale. We repeated the agreement analysis (correlating human-coded body position prevalence with model-predicted prevalence) after dividing the distal comparison period into nine 10-minute bins (marked by vertical dashed lines in Figure \@ref(fig:exemplar-timeline)). Infants had varying numbers of 10-minute bins depending on how much time they were awake and on camera. Bins were included only if there was > 7 minutes of usable data.

@@

After the distal comparison video ended, caregivers were instructed to keep the sensors on their infants for the remainder of the day until infants went to bed, removing the sensors for naps, diaper changes, and trips out of the house. The first two sets of results show that accuracy was consistently high across the proximal and distal recordings, providing confidence that predictions over the remainder of the day would continue to be accurate after the video recordings stopped. This leads to two final questions---how successfully did recordings capture infants' entire day, and how suitable were day-long recordings for capturing individual differences in body position?

@@

Among participants who wore the sensors from morning to bedtime, the majority of the time during the day either resulted in usable body position data or was excluded due to caregiver-reported naps (gray periods in Figure \@ref(fig:timelines)). The total length of the recording period ranged from `r apa_num(min(coverage$total))`-`r apa_num(max(coverage$total))` hours with a median of `r apa_num(median(coverage$total))` hours. Nap times reported during the recording period ranged from `r apa_num(min(coverage$naptime))`-`r apa_num(max(coverage$naptime))` hours with a median of `r apa_num(median(coverage$naptime))` hours. Body position data was available to describe `r apa_num(min(coverage$prop_accounted)*100, digits = 1)`-`r apa_num(max(coverage$prop_accounted)*100, digits = 1)`% of the awake portion of the recording period (median = `r apa_num(median(coverage$prop_accounted)*100, digits = 1)`%). Younger infant \#14 had the least portion of the waking day accounted for; the family left the house to run errands during the majority of the recording period, and the infant napped during much of the remaining time at home. Overall, the recordings produced a median of `r apa_num(median(coverage$hours_recorded))` hours of motion data, with the entire dataset totaling `r apa_num(sum(coverage$hours_recorded))` hours. Even the shortest full recording (`r apa_num(min(coverage$hours_recorded))` hours) exceeded the longest observational sessions in past work in which an experimenter operated a camera.

@@

In the last analysis, we demonstrate how the full-day recordings can reveal individual differences in body position according to age, replicating age differences revealed in past work [@Survey]. Table \@ref(tab:agetable) summarizes the percent of time that younger and older infants spent in each body position out of their awake samples, as predicted by group and individual models. Past work using sparse EMA surveys across the day [@Survey] found that from 3-12 months, infants spent less time held, reclined, and supine but spent increasingly more time sitting and upright. Prone time changed little. The most straightforward comparisons between the two investigations are upright and prone, because they were defined identically. In @Survey, upright time was 0.6-5.5% of the time for infants 3-6 months, increasing to 22.0% at 12 months. Results from the current study were similar: younger infants' upright time estimated from full-day group models was `r pos_sum_lookup["Upright", "Younger_Group"]`% compared with `r pos_sum_lookup["Upright", "Older_Group"]`% for the older group. In @Survey, prone time was 2.9-9.2% of the time for infants 3-6 months, and was 7.2% at 12 months. In the current study, prone time was `r pos_sum_lookup["Prone", "Younger_Group"]`% for younger infants and `r pos_sum_lookup["Prone", "Older_Group"]`% for older infants. Held, sitting, and supine are more difficult to compare because prior work [@Survey] considered held as any time that infants were in caregivers' arms or on their laps, however, we counted laying or sitting on caregivers' laps as supine and sitting, respectively. Further complicating the comparison, @Survey included a reclined category that we would classify as either supine or sitting (whichever it was closer to). Although the numbers are less directly comparable, our results are similar in showing that held and supine time was less for older infants than younger infants (Table \@ref(tab:agetable)). Sitting was greater for older infants (`r pos_sum_lookup["Sitting", "Older_Group"]`%) compared with younger infants (`r pos_sum_lookup["Sitting", "Younger_Group"]`%). Moreover, sitting was the most frequent body position followed by upright in 12-month-olds [@Survey], just as we observed in the current study. Overall, the estimates from the current investigation are consistent with age differences found using a different methodology.

@@

To put this in context, the reliability of human coders on the body position code was ~90%, putting model-predicted accuracy on a similar level to human coders. Acceptable inter-rater reliability in developmental psychology studies of motor behavior is greater than 80%; both group and individual model results exceed that standard for most recorded sessions. Another comparison we can make is to the accuracy of the LENA&reg; device's automated metrics. Correlations between human and model frequencies were *r* = .698 for adult word count and *r* = .649 for child word count [@CristiaLavechin2020]. Even including outliers, correlations between human and model-predicted body position were *r* = `r group_agree_cor` (group) and *r* = `r split_agree_cor` (individual), stronger than the LENA&reg; automated counts. Note, however, that the LENA&reg; validation study used independently-collected data, whereas we are measuring performance for infants whose earlier data comprised the training set. More work is needed to know how well group models would predict a different set of infants' behavior in a different environmental and/or cultural context. 
<!-- Reliability standard? -->

In their validation study, @CristiaLavechin2020 make the recommendation that researchers using LENA&reg; manually annotate a portion of data collected for their particular use case to ensure that automated LENA&reg; metrics are reliable for the participants and environment pertinent to the research question. Similarly, collecting some video data for each participant in future long-form motor behavior recordings provides a way to check reliability and identify errant predictions. Although predictions were accurate for most sessions, we noted two outliers for whom model predictions failed to match human annotations. Fortunately, we could reference video recordings to understand why the errors occurred. Idiosyncrasies in activities and device use make it challenging to decide *a priori* whether odd looking classifications are possible. For example, younger infant \#8 on Figure \@ref(fig:timelines) spent an unusual amount of time upright at an age where infants cannot yet stand. Inspecting the video revealed that the infant spent long periods of time suspended in a jumper that supported their body in an upright position. Although it is not feasible to collect large amounts of video data to check model predictions, interviewing caregivers about common activities and devices may provide a way to understand unexpected predictions.

@@

Aside from differences in validity, researchers may favor one modeling approach based on logistical concerns. Group models have several major advantages. First, by definition they are trained on more data, which might make predictions more consistent. Second, group models remove the need to get optimal training data from each participant. It can be difficult to elicit every behavior of interest in each infant; individual models require that every behavior that will be later predicted was displayed by each individual infant. In fact, group models remove the need to get *any* training data for a particular infant, meaning that they can be applied to any infant regardless of whether they have training data. Finally, applying a group model means that individual differences between infants are due to their behavior being classified in a consistent way, rather than having a separate set of rules for predicting each individual. However, given the variety in infant behavior, individual models are more agile in that they can capture the unique and unexpected. Likewise, individual models can be an excellent choice for rapid prototyping and pilot testing. A researcher can get proof-of-concept data from a single participant or a new kind of classification scheme (i.e., locomotion, restrain in device) without annotating an entire sample. Individual models may also be applied in clinical cases where infants vary widely in their motor abilities and/or use of assistive devices.

Beyond group and individual models, there were many degrees of freedom in our choice of how best to model body position. For the sake of brevity, we report the best approach, not every possible variation in modeling. However, our openly shared data and code allow researcher to experiment with different sets of motion features, different sets of training data, different machine learning algorithms, and different hyperparameter values. As in past work [@AiraksinenRasanen2020], performance is best when using all four sensors and degrades when using any subset of two (i.e., hips only, left leg only). We tested but did not report models that used either only accelerometer data and only gyroscope data; in both cases, performance is degraded compared with models that use both types of motion. We also tried different machine learning algorithms (XGBoost, Tabnet), but found that random forest models performed the best. Finally, we tried modeling each age group separately, but found no benefit to performance by tailoring to the particular age group.
<!-- DATA CITE -->

@@

Long-form recordings of motor behavior bring about new research possibilities. Although past work using video [@Karasik2011; @ChenSchneider2022; @HerzbergFletcher2021] produced real-time data, such data were limited to a relatively small part of the infants' day. And although survey methods [@Survey] could capture moments scattered throughout the day, they do not produce real-time data. The combination of real-time, full-day data about infant motor behavior is unprecedented and offers new opportunities for understanding infants' everyday experiences. Collecting dense data over the entire variety of daily experiences helps to more accurately measure infants' experiences in aggregate (as in Figure \@ref(fig:age)) without biasing results from a particular type of activity (e.g., play). Long-form recordings also have the potential to measure clinically-relevant outcomes, such as whether a physical therapy intervention to improve infant sitting skill leads to increases in everyday sitting time.

In the past decade, developmental scientists have discovered that distributional information about infants' experiences matter for how infants learn [@ClerkinHart2017; @KachergisYu2017; @RazAbney2019]. Skewed distributions---those that favor the repetition of a small subset of experiences---facilitate learning. For example, recordings from wearable head cameras in the home indicate that infants see a small subset of objects with high frequency [@ClerkinHart2017], but most objects are seen infrequently. Heterogeneity is also found in how experiences are distributed in time---burstiness and clustering are seen in infants' daily experiences seeing objects and hearing music. Long-form motor recordings provide a novel opportunity to measure the temporal structure of motor experiences that can be seen in Figure \@ref(fig:timelines). How skewed are infants' experiences with different motor behaviors, and how are they clustered in time? Linking the distributional structure of a particular motor experience in a real day, such as time spent sitting, might be predicted by sitting skill and/or might predict infants' future sitting ability. 
<!-- CITES FOR BURSTINESS CASILLAS FAUSEY -->

@@

What underpins these novel research applications is that body position classifications can be applied automatically at scale. Annotating a video corpus of infant behavior from a moderately sized sample---such as 40 infants recorded for 2 hours each in 2 separate visits [@HerzbergFletcher2021]---is incredibly labor intensive. Annotating a larger, more representative sample of *hundreds of infants* while simultaneously scoring *full-day data* would be prohibitive. With sufficiently accurate group models, researchers could annotate a moderate video sample to then apply the model to hundreds of full-day recordings. The added cost of more recording time is low, meaning that future studies could sample across multiple days of behavior to better understand intra- versus inter-day variability. Prior diary methods show that infants inconsistently display new motor skills on a day-to-day basis [@AdolphRobinson2008; @AdolphRobinson2011]. Multi-day recordings could further uncover how infants' experiences vary over timescales from seconds to days.

<!-- More easily shareable data (non-identifiable, text based) -->

>>>>

Three online repositories contain openly shared data, materials, and analysis code. A Databrary repository (https://nyu.databrary.org/volume/1580) includes an exemplar participant's recording session, with the raw video data files, the Datavyu annotations of those video files, a log file with machine-readable synchronization points and nap/diaper change times, and accelerometer and gyroscope data for each of the 4 sensors. A CodeOcean capsule (LINK) hosts an R environment with the exemplar participant's data and source code to: 1) synchronize IMU and video annotations, 2) calculate windowed motion features for their data, and 3) train and test the body position classifier using an "individual model". The chosen participant's classification accuracy was at the median level. Because of the overall size of the full dataset and the computational power/time required to synchronize and create windowed datasets for each session, it would not be feasible to create a reproducible environment to process all 34 sessions. However, in a second CodeOcean capsule (LINK) we share the entire result: The dataset of windowed motion features with corresponding body position codes used to validate the method (described in the results below), along with the processed intermediary files used for each set of analyses. This reproducible manuscript created in *papaja* [@papaja] can be executed in the CodeOcean environment for full computational transparency, and can be downloaded from Github (LINK).
<!-- FINAL LINKS -->

@@

We report three sets of results based on `r length(unique(ppts$id_uni))` full-day testing sessions resulting in a total of `r round(sum(duration$duration_hours))` hours of movement recording. First, we focus on the accuracy of group and individual models trained and validated from data during the proximal comparison period. This period is unique in having the highest degree of temporal synchrony between video and motion data, allowing us to assess the accuracy of individual body position events. Second, we applied models trained from data during the proximal comparison to predict infant behavior during the distal comparison, when coarser synchrony between video and motion data was available. This distal comparison is critical because it provides the first-ever test of how well infant body classification models predict behavior that is variable, unrestricted, and unsupervised. Data from the first two sets of analyses confirm that body position classification models are accurate across a long period of time, suggesting that data from the entire day will be valid. In the third set of results we examine the data quality of full-day recordings and test whether full-day recordings capture well-established age differences in body position experiences.

@@

The first set of analyses use data from the proximal comparison to determine the "best case" accuracy of the models, training and testing on similar types of data. The high degree of temporal synchronization between video and motion data during this period makes it possible to link human-coded body position annotations to each 4-s window of motion data, providing ground truth data for model training and testing. As in past work [@FranchakScott2021], we compared two types of models: *group models* and *individual models*. To assess the accuracy of body position classification for each recording session, we reserved the (temporally) last 25% of a session's proximal comparison data as the testing set. The testing set was never used as training data. The two types of models differed based on what data were used during training, but the same random forest modeling procedure was used in both cases. A group model was trained for each session using an aggregated dataset containing the first 75% of all **other** sessions' proximal comparison data. This leave-one-out cross-validation tested the generalization of the model to a recording session that was not used at all in the training set. In contrast, individual models were trained using the first 75% of a recording session, then tested against the (temporally) last 25% of the same session. The individual model tests whether earlier training data generalize to later testing data within an individual participant's recording. In both cases, the testing set was the same, allowing for direct comparison between the two modeling approaches. Figure \@ref(fig:metrics), Table \@ref(tab:metrics-table), and Table \@ref(tab:metrics-by-class) summarize the performance of group and individual models using standard metrics for classification.

@@

Overall accuracy (Figure \@ref(fig:metrics)A) represents the proportion of 4-s windows in the testing set in which the model prediction matched the human annotation of body position. Overall accuracy for group models  (*M* = `r metrics_lookup["Overall Accuracy", "mean_Group"]`) was slightly lower than accuracy for individual models (*M* = `r metrics_lookup["Overall Accuracy", "mean_Individual"]`). Although overall accuracy from our semi-supervised, in-home data collection did not match the near-perfect accuracy  (.95-.98) found in prior in-lab studies [@FranchakScott2021; @AiraksinenRasanen2020], both models approached the level of agreement found between two human coders (*M* = .906). Most likely, lower accuracy in the current study results from the more variable and complex behavior observed in a a semi-supervised setting rather than from a difference in the quality of the classification model; although the first part of the proximal period was guided by the experimenter during a brief phone call, the remainder of the proximal period included natural behavior. Visual inspection of Figure \@ref(fig:metrics)A shows that accuracy values were heavily skewed, with many approaching perfect accuracy but a few sessions with very poor accuracy. Looking at the median performance suggests that the difference between models was not considerable for the typical participant (group median accuracy = `r metrics_lookup["Overall Accuracy", "median_Group"]`; individual median accuracy = `r metrics_lookup["Overall Accuracy", "median_Individual"]`). Indeed, it is notable that the worst-case accuracy from group models (minimum = `r metrics_lookup["Overall Accuracy", "min_Group"]`) was far lower than in individual models (minimum = `r metrics_lookup["Overall Accuracy", "min_Individual"]`). Possibly, individual models accounted for idiosyncrasies in behavior or inconsistencies in sensor placement for those sessions that group models could not generalize to.

### Cohen's Kappa 
Strong overall accuracy can be misleading when the relative frequency of different classes is unbalanced. For example, most infants spent more time sitting than prone, so good performance for sitting classification could mask poorer performance in prone classification. Accordingly, we report Cohen's kappa, a commonly-used metric that penalizes missing rare events (Figure \@ref(fig:metrics)B), and we provide classification metrics for each individual body position (Table \@ref(tab:metrics-by-class)) to account for imbalance in body position rates within and between individuals. Similar to overall accuracy, kappa values were strong for both model types with group kappas (*M* = `r metrics_lookup["Kappa", "mean_Group"]`) somewhat worse compared with individual kappas (*M* = `r metrics_lookup["Kappa", "mean_Individual"]`). Guidelines for interpreting kappa statistics [@LandisKoch1977] consider 0.81–1.00 "Almost Perfect,” 0.61–0.80 “Substantial,” 0.41–0.60 “Moderate,” 0.21– 0.40 “Fair,” and 0–0.20 “Slight to Poor”, indicating that agreement for most group and individual model predictions fell in the Substantial to Almost Perfect range.

@@

Figure \@ref(fig:part2overall) and Table \@ref(tab:pt2overalltable) summarize the overall agreement during the distal comparison period. For each session, we calculated the actual time spent in each body position (out of times the infant was visible on camera and awake) using human annotated body position (x-axis on Figure \@ref(fig:part2overall)). Predicted time was calculated the same way for group and individual model predictions, omitting the off-camera and nap periods to make a direct comparison. Overall---across participants and across body position classes---agreement was strong: The correlation between group model predictions and human-coded time was *r* = `r group_agree_cor`, and the correlation between individual model predictions and human-coded time was *r* = `r split_agree_cor`. As in the proximal comparison, agreement varied somewhat between body positions; in particular, agreement for held was poor. Unlike in the proximal period, some body positions were better predicted by group models and others by individual models.

Visual inspection of Figure \@ref(fig:part2overall) indicated two extreme outliers, which we marked by a gray square and a gray diamond. We closely investigated each of the two outliers to understand why their agreement was below the standard of the other sessions. The "gray square" outlier had significant confusion between sitting and supine classification. Reviewing the video indicated that this participant spent a long period of time in a seating device that was reclined almost exactly at 45 degrees, making it difficult to determine if the infant was sitting or supine. The infant also spent a long time in the mother's arms in an ambiguous supine/sitting position that was hard to classify. This participant's proximal accuracy was also poor because similar ambiguities appeared during the initial recording. In contrast, the "gray diamond" outlier had strong proximal accuracy, with confusion only arising in the distal period between upright and held categories. Reviewing the video showed that all disagreements occurred when the infant was in a baby walker; human coders scored this as "upright" but the models predicted it as "held". Most likely, the infant's movements in the baby walker were more similar to how a baby moved while carried, and unlike how most infants moved while walking upright.

@@

Collecting real-time data that can be synchronized to other sources is a key goal in our development of long-form body position classification. Although overall aggregate agreement in the distal comparison was strong, it is important to show that similarly strong agreement is found within a shorter timescale. We repeated the agreement analysis (correlating human-coded body position time with model-predicted time) after dividing the distal comparison period into nine 10-minute bins (marked by vertical dashed lines in Figure \@ref(fig:exemplar-timeline)). Infants had varying numbers of 10-minute bins depending on how much time they were awake and on camera. Bins were included only if there was > 7 minutes of usable data.

@@

After the distal comparison video ended, caregivers were instructed to keep the sensors on their infants for the remainder of the day until infants went to bed, removing the sensors for naps, diaper changes, and trips out of the house. The first two sets of results show that accuracy was consistently high across the proximal and distal recordings, providing confidence that predictions over the remainder of the day would continue to be accurate after the video recordings stopped. This leads to two final questions---how successfully did recordings capture infants' entire day, and how well do findings from full-day classification converge with findings that use other methodologies?

@@

Among participants who wore the sensors from morning to bedtime, the majority of the time during the day either resulted in usable body position data or was excluded due to caregiver-reported naps (gray periods in Figure \@ref(fig:timelines)). The total length of the recording period ranged from `r apa_num(min(coverage$total))`-`r apa_num(max(coverage$total))` hours with a median of `r apa_num(median(coverage$total))` hours. Nap times reported during the recording period ranged from `r apa_num(min(coverage$naptime))`-`r apa_num(max(coverage$naptime))` hours with a median of `r apa_num(median(coverage$naptime))` hours. Body position data were available to describe `r apa_num(min(coverage$prop_accounted)*100, digits = 1)`-`r apa_num(max(coverage$prop_accounted)*100, digits = 1)`% of the awake portion of the recording period (median = `r apa_num(median(coverage$prop_accounted)*100, digits = 1)`%). Younger infant \#14 had the least portion of the waking day accounted for; the family left the house to run errands during the majority of the recording period, and the infant napped during much of the remaining time at home. Overall, the recordings produced a median of `r apa_num(median(coverage$hours_recorded))` hours of motion data, with the entire dataset totaling `r apa_num(sum(coverage$hours_recorded))` hours. Even the shortest full recording (`r apa_num(min(coverage$hours_recorded))` hours) exceeded the longest observational sessions in past work in which an experimenter operated a camera.

@@

In order to show the convergent validity of the classification method, we demonstrate how the full-day recordings can reveal individual differences in body position according to age, replicating age differences revealed in past work [@Survey]. Table \@ref(tab:agetable) summarizes the percent of time that younger and older infants spent in each body position out of their awake samples, as predicted by group and individual models. Past work using sparse EMA surveys across the day [@Survey] found that from 3-12 months, infants spent less time held, reclined, and supine but spent increasingly more time sitting and upright. Prone time changed little. The most straightforward comparisons between the two investigations are upright and prone, because they were defined identically. In @Survey, upright time was 0.6-5.5% of the time for infants 3-6 months, increasing to 22.0% at 12 months. Results from the current study were similar: younger infants' upright time estimated from full-day group models was `r pos_sum_lookup["Upright", "Younger_Group"]`% compared with `r pos_sum_lookup["Upright", "Older_Group"]`% for the older group. In @Survey, prone time was 2.9-9.2% of the time for infants 3-6 months, and was 7.2% at 12 months. In the current study, prone time was `r pos_sum_lookup["Prone", "Younger_Group"]`% for younger infants and `r pos_sum_lookup["Prone", "Older_Group"]`% for older infants. Held, sitting, and supine are more difficult to compare because prior work [@Survey] considered held as any time that infants were in caregivers' arms or on their laps, however, we counted laying or sitting on caregivers' laps as supine and sitting, respectively. Further complicating the comparison, @Survey included a reclined category that we would classify as either supine or sitting (whichever it was closer to). Although the numbers are less directly comparable, our results are similar in showing that held and supine time was less for older infants than younger infants (Table \@ref(tab:agetable)). Sitting was greater for older infants (`r pos_sum_lookup["Sitting", "Older_Group"]`%) compared with younger infants (`r pos_sum_lookup["Sitting", "Younger_Group"]`%). Moreover, sitting was the most frequent body position followed by upright in 12-month-olds [@Survey], just as we observed in the current study. Overall, the estimates from the current investigation are consistent with age differences found using a different methodology.

@@

To put this in context, the reliability of human coders on the body position code was ~90%, putting model-predicted accuracy on a similar level to human coders. Another comparison we can make is to the accuracy of the LENA&reg; device's automated metrics. Correlations between human and model frequencies were *r* = .698 for adult word count and *r* = .649 for child word count [@CristiaLavechin2020]. Even including outliers, correlations between human and model-predicted body position were *r* = `r group_agree_cor` (group) and *r* = `r split_agree_cor` (individual), stronger than the LENA&reg; automated counts. Note, however, that the LENA&reg; validation study used independently-collected data, whereas we are measuring performance for infants whose earlier data comprised the training set. More work is needed to know how well group models would predict a different set of infants' behavior in a different environmental and/or cultural context.

In their validation study, @CristiaLavechin2020 make the recommendation that researchers using LENA&reg; manually annotate a portion of data collected for their particular use case to ensure that automated LENA&reg; metrics are reliable for the participants and environment pertinent to the research question. Similarly, collecting some video data for each participant in future long-form motor behavior recordings provides a way to check reliability and identify errant predictions. Although predictions were accurate for most sessions, we noted two outliers for whom model predictions failed to match human annotations. Fortunately, we could reference video recordings to understand why the errors occurred. Idiosyncrasies in activities and device use make it challenging to decide *a priori* whether "odd" classifications (from the expectation of the researcher) are possible. For example, younger infant \#8 on Figure \@ref(fig:timelines) spent an unusual amount of time upright at an age where infants cannot yet stand. Inspecting the video revealed that the infant spent long periods of time suspended in a jumper that supported their body in an upright position. Although it is not feasible to collect large amounts of video data to check model predictions, interviewing caregivers about common activities and devices may provide a way to understand unexpected predictions.

@@

Aside from differences in validity, researchers may favor one modeling approach based on logistical concerns. Group models have several major advantages. First, by definition they are trained on more data, which might make predictions more consistent. Second, group models remove the need to get optimal training data from each participant. It can be difficult to elicit every behavior of interest in each infant; individual models require that every behavior that will be later predicted was displayed by each individual infant. In fact, group models remove the need to get *any* training data for a particular infant, meaning that they can be applied to any infant regardless of whether they have training data. Finally, applying a group model means that individual differences between infants are due to their behavior being classified in a consistent way, rather than having a separate set of rules for predicting each individual. However, given the variety in infant behavior, individual models are more agile in that they can capture the unique and unexpected. Likewise, individual models can be an excellent choice for rapid prototyping and pilot testing. A researcher can get proof-of-concept data from a single participant or a new kind of classification scheme (i.e., locomotion, restraint in device) without annotating an entire sample. Individual models may also be applied in clinical cases where infants vary widely in their motor abilities and/or use of assistive devices.

Beyond group and individual models, there were many degrees of freedom in our choice of how best to model body position. For the sake of brevity, we chose to report the best approach, not every possible variation in modeling. However, our openly shared data and code allow researcher to experiment with different sets of motion features, different sets of training data, different machine learning algorithms, and different hyperparameter values. As in past work [@AiraksinenRasanen2020], performance was best when using all four sensors and degraded when using any subset of two (i.e., hips only, left leg only). We tested but did not report models that used either only accelerometer data and only gyroscope data; in both cases, performance was degraded compared with models that use both types of motion. We also tried different machine learning algorithms (XGBoost, Tabnet), but found that random forest models performed the best. Finally, we tried modeling each age group separately, but found no benefit to performance by tailoring to the particular age group.

@@

Long-form recordings of motor behavior bring about new research possibilities. Although past work using video [@Karasik2011; @ChenSchneider2022; @HerzbergFletcher2021] produced real-time data, such data were limited to a relatively small part of the infants' day. And although survey methods [@Survey] could capture moments scattered throughout the day, they do not produce real-time data. The combination of real-time, full-day data about infant motor behavior is unprecedented and offers new opportunities for understanding infants' everyday experiences. Collecting dense data over the entire variety of daily experiences helps to more accurately measure infants' experiences in aggregate (as in Figure \@ref(fig:age)) without biasing results from a particular type of activity (e.g., play). Long-form recordings also have the potential to measure clinically-relevant outcomes for infants with motor delay or other pediatric concerns.

In the past decade, developmental scientists have discovered that distributional information about infants' experiences matter for how infants learn [@ClerkinHart2017; @KachergisYu2017; @RazAbney2019]. Skewed distributions---those that favor the repetition of a small subset of experiences---facilitate learning. For example, recordings from wearable head cameras in the home indicate that infants see a small subset of objects with high frequency [@ClerkinHart2017], but most objects are seen infrequently. Heterogeneity is also found in how experiences are distributed in time---burstiness and clustering are seen in infants' daily experiences seeing objects and hearing music [@MendozaFausey2022; @CasillasElliott2021]. Long-form motor recordings provide a novel opportunity to measure the temporal structure in the sequences of body positions seen in Figure \@ref(fig:timelines). How skewed are infants' experiences with different motor behaviors, and how are they clustered in time? Linking the distributional structure of a particular motor experience in a real day, such as time spent sitting, might be predicted by sitting skill and/or might predict infants' future sitting ability.

@@

What underpins these novel research applications is that body position classifications can be applied automatically at scale. Annotating a video corpus of infant behavior from a moderately sized sample---such as 40 infants recorded for 2 hours each in 2 separate visits [@HerzbergFletcher2021]---is incredibly labor intensive. Annotating a larger, more representative sample of *hundreds of infants* while simultaneously scoring *full-day data* would be prohibitive. With sufficiently accurate group models, researchers could annotate a moderately-sized video corpus to then apply the model to hundreds of full-day recordings. Unlike with video, the added cost of more IMU sensor recording time is low, meaning that future studies could sample across multiple days of behavior to better understand intra- versus inter-day variability. Prior diary methods show that infants inconsistently display new motor skills on a day-to-day basis [@AdolphRobinson2008; @AdolphRobinson2011]. Multi-day recordings could further uncover how infants' experiences vary over timescales from seconds to days. 
<!-- More easily shareable data (non-identifiable, text based) -->
